draft_yma

### 2021-12-16 draft_yma.Rmd
all code should be here, discuss tmr

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
```

@@

```{r, message=FALSE}
cdi = read_csv("cdi.csv") %>%
  janitor::clean_names() %>%
  mutate(crm_1000 = 1000*(crimes/pop),
         pdocs_1000 = 1000*(docs/pop),
         pbeds_1000 = 1000*(beds/pop),
         density_pop = pop/area,
         # convert region to factors and recoded them accordingly 
         region = factor(region, levels = 1:4,
                    labels = c("northeast", "north central", "south", "west"))) %>% select(-c(docs,beds)) %>% view()

@@

### States with unusual rates

group by state
```{r}
cdi_by_state = cdi %>% group_by(state) %>% 
  summarise(state_pop = sum(pop),
            state_crimes = sum(crimes)) %>% 
  mutate(state_CRM_1000 = 1000*state_crimes/state_pop)
```

Not sure what counts as out lier, this is what the other group did.

@@

This shows no outlier on the state level tho...

```{r}
upper = quantile(cdi_by_state$state_CRM_1000, 0.75)
lower = quantile(cdi_by_state$state_CRM_1000, 0.25)
IQR = upper - lower
cdi_by_state %>% 
  filter(state_CRM_1000 > upper + 1.5*IQR,
         state_CRM_1000 > lower - 1.5*IQR) %>% 
  dplyr::select(state, state_CRM_1000) %>%
  knitr::kable(digits = 2)
boxplot(cdi_by_state$state_CRM_1000, main = 'State Crime Rate per 1000 people')
```

**we probably need to plot region as well? not sure**

@@

```{r}
trans_fit = lm(crm_1000^0.5 ~ .,data = cdi_model)
summary(trans_fit)

@@

lambda_trans = MASS::boxcox(trans_fit)
optlam_trans = lambda_trans$x[which.max(lambda_trans$y)]
lambda_trans

cdi_trans = cdi_model %>% mutate(crm_1000_sqr = crm_1000^0.5) %>% select(-c(crm_1000))
```

@@

lm(formula = crm_1000^0.5 ~ pop + pop18 + hsgrad + bagrad + poverty + 
    pcincome + region + pbeds_1000 + density_pop, data = cdi_model)

@@

```{r}
fit_forward = step(trans_fit, direction='forward')
fit_forward

@@

lm(formula = crm_1000^0.5 ~ pop + pop18 + pop65 + hsgrad + bagrad + 
    poverty + unemp + pcincome + region + pdocs_1000 + pbeds_1000 + 
    density_pop, data = cdi_model)

@@

@@

```{r}
olsrr::ols_mallows_cp(fit_back,trans_fit)
olsrr::ols_mallows_cp(fit_forward,trans_fit)
olsrr::ols_mallows_cp(fit_both,trans_fit)
```

@@

```{r}
summary(fit_back) # 0.5469
summary(fit_forward) #0.5457
summary(fit_both) #5469
```

### BIC

```{r}
library(flexmix)
BIC(fit_back)
BIC(fit_forward)
BIC(fit_both)
```

## Cp

didn't use any of this!
```{r, include=FALSE}
# printing the 2 best models of each size. For example, the first two lines: print the best 2 models that have 2 variables (including intercept)
library(leaps)
Cp_b = leaps(x = model.matrix(trans_fit)[,-1],
      y = cdi_trans$crm_1000_sqr,
      nbest = 2,
      method = "Cp")

mod_b = Cp_b$Cp[which(Cp_b$Cp <= Cp_b$size)]
```

size is the number of parameters in the model, Cp is Cp

The plot shows the model that satisfy the criteria, number indicates the number of parameters it included. in this case, the 1st to the last.

**forward**
```{r, include=FALSE}
Cp_f = leaps(x = model.matrix(fit_forward)[,-1],
      y = cdi_trans$crm_1000_sqr,
      nbest = 2,
      method = "Cp")

mod_f = Cp_f$Cp[which(Cp_f$Cp <= Cp_f$size)]
```

So, multiple model satisfies the Cp criteria here.

Did not use!!
**backward**
```{r, include=FALSE}
adjr_b = leaps(x = model.matrix(fit_back)[,-1],
      y = cdi_trans$crm_1000_sqr,
      nbest = 2,
      method = "adjr2")

adjr_b

adjr_b%>% 
  faraway::maxadjr(10)
```
**forward**
```{r, include=FALSE}
leaps(x = model.matrix(fit_forward)[,-1],
      y = cdi_trans$crm_1000_sqr,
      nbest = 2,
      method = "adjr2") %>% faraway::maxadjr(10)
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(leaps)
library(readr)
```

@@

```{r, message=FALSE}
cdi = read_csv("cdi.csv") %>%
  janitor::clean_names() %>%
  mutate(crm_1000 = 1000*(crimes/pop),
         pdocs_1000 = 1000*(docs/pop),
         pbeds_1000 = 1000*(beds/pop),
         density_pop = pop/area,
         # convert region to factors and recoded them accordingly 
         region = factor(region, levels = 1:4,
                    labels = c("northeast", "northcentral", "south", "west"))) %>% select(-c(docs,beds))

@@

### Counties with unusual rates

@@

@@

### New tibble with transformed crm_1000 -> crm_1000_sqr
```{r}
cdi_trans = cdi_model %>% mutate(crm_1000_sqr = crm_1000^0.5) %>% select(-c(crm_1000))
```

```{r}
trans_fit = lm(crm_1000_sqr ~ .,data = cdi_trans)
summary(trans_fit)

@@

lambda_trans = MASS::boxcox(trans_fit)
optlam_trans = lambda_trans$x[which.max(lambda_trans$y)]
optlam_trans
```

### Counties with unusual rates

```{r}
upper = quantile(cdi_trans$crm_1000_sqr, 0.75)
lower = quantile(cdi_trans$crm_1000_sqr, 0.25)
IQR = upper - lower

cdi_trans %>% 
  filter(crm_1000_sqr > upper + 1.5*IQR,
         crm_1000_sqr > lower - 1.5*IQR) %>% 
  dplyr::select(crm_1000_sqr) %>% 
  mutate(cty = c("Kings", "St._Loui")) %>% 
  knitr::kable(digits = 2)

# cdi %>% filter(pop == 2300664 | pop == 396685) %>% select(cty,pop)
```

@@

Backward: crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + pcincome + region + pbeds_1000 + density_pop

@@

```{r}
fit_forward = step(trans_fit, direction="forward")
fit_forward

@@

lm(formula = crm_1000_sqr ~ pop + pop18 + pop65 + hsgrad + bagrad + 
    poverty + unemp + pcincome + region + pdocs_1000 + pbeds_1000 + 
    density_pop, data = cdi_trans)

@@

## Criterion Based: Adjr2 and Cp

```{r}
# printing the 2 best models of each size. For example, the first two lines: print the best 2 models that have 2 variables (including intercept)

models <- regsubsets(crm_1000_sqr ~., data = cdi_trans,nvmax = 12)
res_sum = summary(models)

par(mfrow=c(1,2))
plot(1:12, res_sum$cp, xlab="No of parameters", ylab="Cp Statistic")
plot(1:12, res_sum$adjr2, xlab="No of parameters", ylab="Adj R2")

mod_s = data.frame(
  Adj.R2 = which.max(res_sum$adjr2),
  CP = which.min(res_sum$cp)
)

get_model_formula <- function(id, object, outcome){
  # get models data
  models <- summary(object)$which[id,-1]
  # Get outcome variable
  #form <- as.formula(object$call[[2]])
  #outcome <- all.vars(form)[1]
  # Get model predictors
  predictors <- names(which(models == TRUE))
  predictors <- paste(predictors, collapse = "+")
  # Build model formula
  as.formula(paste0(outcome, "~", predictors))
}
```

Model chosen by Adj $R^2$:

```{r}
get_model_formula(12, models, "crm_1000_sqr")
```

```{r}
fit_adjr = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + region+ pbeds_1000 + density_pop, data = cdi_trans)

olsrr::ols_plot_resid_fit(fit_adjr)
olsrr::ols_plot_resid_qq(fit_adjr)
MASS::boxcox(fit_adjr)
```

crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + region+ pbeds_1000 + density_pop

Model chosen by Cp: Same with backward

```{r}
get_model_formula(11, models, "crm_1000_sqr")
```

crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + pcincome + region+ pbeds_1000 + density_pop

```{r}
fit_Cp = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + pcincome + region+ pbeds_1000 + density_pop, data=cdi_trans)

olsrr::ols_plot_resid_fit(fit_Cp)
olsrr::ols_plot_resid_qq(fit_Cp)
MASS::boxcox(fit_Cp)
```

@@

```{r}
olsrr::ols_mallows_cp(fit_back,trans_fit)
olsrr::ols_mallows_cp(fit_forward,trans_fit)
olsrr::ols_mallows_cp(fit_both,trans_fit)
olsrr::ols_mallows_cp(fit_adjr,trans_fit)
olsrr::ols_mallows_cp(fit_Cp,trans_fit)
```

@@

```{r}
summary(fit_back) # 0.5469
summary(fit_forward) #0.5457
summary(fit_both) #0.5469
summary(fit_adjr) #0.5474
summary(fit_Cp) #0.5469
```

#### BIC

```{r}
library(flexmix)
BIC(fit_back) #1453.527
BIC(fit_forward) #1469.908
BIC(fit_both) #1453.527
BIC(fit_adjr) # 1458.104
BIC(fit_Cp) #1453.527
```

#### MSE

```{r}
back_summ =summary(fit_back) #1.331
mean(back_summ$residuals^2)

forward_summ =summary(fit_forward) #1.325
mean(forward_summ$residuals^2)

both_summ =summary(fit_both) #1.331
mean(both_summ$residuals^2)

adjr_summ =summary(fit_adjr) #1.326
mean(adjr_summ$residuals^2)

Cp_summ =summary(fit_Cp) #1.331
mean(Cp_summ$residuals^2)
```

## Check Collinearity

```{r}
performance::check_collinearity(fit_back)
performance::check_collinearity(fit_forward)
performance::check_collinearity(fit_adjr)
performance::check_collinearity(fit_Cp)
```

delete vif larger than 10. Non of them are greater than 10. So all of the models are good, don't have to worry about collinearity.

```{r}
par(mfrow=c(2,2))
plot(fit_back)
```

## Checking to Outliers and Influential Points

### Back

#### Old Method
```{r}
# residuals vs leverage plot
plot(fit_back, which = 4)
#remove outliers
cdi_trans_without = cdi_trans[-c(1, 6, 53),]
fit_without = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + 
    pcincome + region + pbeds_1000 + density_pop, data = cdi_trans_without)
summary(fit_without)
```

```{r}
par(mfrow=c(2,2))
plot(fit_without)
```

### Threshold: $D_{i}>\frac{4}{n} = \frac{4}{440} = 0.009$
```{r}
cook = olsrr::ols_plot_cooksd_bar(fit_back, print_plot = TRUE)
normal = cook$data$color == "normal"
outlier = cook$data$color == "outlier"
```

```{r}
outlier_df = cdi[outlier,]
```

This gives us 21 outliers. We should consider if we want to take them all out

### Build model without outlier

Here's what happens when we take them all out.

```{r}
cdi_trans_without = cdi_trans[normal,]
```

Built the without model with **fit_back**

```{r}
fit_without = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + pcincome + region + pbeds_1000 + density_pop, data = cdi_trans_without)

summary(fit_without)

# Residual performance
par(mfrow=c(2,2))
olsrr::ols_plot_resid_fit(fit_without)
olsrr::ols_plot_resid_qq(fit_without)
MASS::boxcox(fit_without)

#Cp -88.45??
olsrr::ols_mallows_cp(fit_without,trans_fit)
#BIC: 1274.534
BIC(fit_without) 
#MSE
without_summ =summary(fit_without) #1.017
mean(without_summ$residuals^2)
#VIF 
performance::check_collinearity(fit_without)
```

Built the without model with **fit_adjr**

```{r}
fit_without2 = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + region+ pbeds_1000 + density_pop, data = cdi_trans_without)

summary(fit_without2)

# Residual performance
par(mfrow=c(2,2))
olsrr::ols_plot_resid_fit(fit_without2)
olsrr::ols_plot_resid_qq(fit_without2)
MASS::boxcox(fit_without2)

#Cp -87.5??
olsrr::ols_mallows_cp(fit_without2,trans_fit)
#BIC: 1279.154
BIC(fit_without2)
#MSE
without2_summ =summary(fit_without2) #1.013
mean(without2_summ$residuals^2)
#VIF 
performance::check_collinearity(fit_without2) 
```

# Model Validation

## Compute RMSE and adjusted R^2 by cross-validation

Using back and adjr model and data without outlier

```{r warning=FALSE}
library(modelr)
set.seed(1)
cv_df = 
  crossv_kfold(cdi_trans_without, k = 5) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

cv_df = 
  cv_df %>%
  mutate(
    fit_back = map(train, ~lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + pcincome + region + pbeds_1000 + density_pop, data = .x)),
    trans_fit = map(train, ~lm(crm_1000_sqr ~ pop + pop18 + pop65+ hsgrad + bagrad + poverty + unemp + pcincome + region+ +pdocs_1000+pbeds_1000 + density_pop, data = .x)),
    fit_adjr = map(train, ~lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + region+ pbeds_1000 + density_pop, data = .x)),
    fit_forward = map(train, ~lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + region+ pbeds_1000 + density_pop, data = .x))) %>% 
  mutate(
    rmse_model2 = map2_dbl(fit_back, test, ~rmse(model = .x, data = .y)),
    rmse_model3 = map2_dbl(trans_fit, test, ~rmse(model = .x, data = .y)),
    rmse_model4 = map2_dbl(fit_adjr, test, ~rmse(model = .x, data = .y)),
    rmse_model5 = map2_dbl(fit_forward, test, ~rmse(model = .x, data = .y))) %>%
  mutate(
    mspe_model2 = map2_dbl(fit_back, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2)),
    mspe_model3 = map2_dbl(trans_fit, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2)),
    mspe_model4 = map2_dbl(fit_adjr, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2)),
    mspe_model5 = map2_dbl(fit_forward, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2))
  ) %>% 
  mutate(
    res_model2 = map(fit_back, broom::glance %>% as.data.frame),
    res_model3 = map(trans_fit, broom::glance %>% as.data.frame),
    res_model4 = map(fit_adjr, broom::glance %>% as.data.frame),
    res_model5 = map(fit_forward, broom::glance %>% as.data.frame)) %>%
  unnest(res_model2, res_model3, res_model4, res_model5) %>%
  dplyr::select(rmse_model2,rmse_model3,rmse_model4,rmse_model5,
                mspe_model2,mspe_model3,mspe_model4,mspe_model5,
                value.adj.r.squared,value.adj.r.squared1,value.adj.r.squared2,value.adj.r.squared3) %>%
  rename(adjR_model2 = value.adj.r.squared,
         adjR_model3 = value.adj.r.squared1,
         adjR_model4 = value.adj.r.squared2,
         adjR_model5 = value.adj.r.squared3)

cv_df %>%
  summarise_each(funs(mean( .,na.rm = TRUE))) %>%
  t()
```

## Plot the violin plot
```{r warning=FALSE}
unnest_cd_df = cv_df %>%
  pivot_longer(rmse_model2:adjR_model5, 
               names_pattern = "(.*)(......)$", 
               names_to = c("limit", "model")) %>% 
  mutate(limit=ifelse(limit=="", "value", limit)) %>%
  pivot_wider(id_cols = model, 
              names_from = limit, 
              values_from = value, 
              names_repair = "check_unique") %>% 
  unnest(c(rmse_, adjR_)) %>%
  rename(rmse = rmse_,
         adjR = adjR_)
unnest_cd_df %>% ggplot(aes(x = model, y = rmse)) + geom_violin() 
unnest_cd_df %>% ggplot(aes(x = model, y = adjR)) + geom_violin() 
```
